---
title: "Project 2 Code: Public Data Visualisation and Mapping with R"
subtitle: "LSE MY472 Portfolio Project"
format:
  html:
    embed-resources: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

Project 2 Code: Public Data Visualisation and Mapping with R

This file contains the full source code for Project 2. The project uses the original Week 4 workflow and adds output-saving code for the portfolio.

1. Set up

```{r}
required_packages <- c("tidyverse", "sf", "scales", "gridExtra")

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
}

library(tidyverse)
library(sf)
library(scales)
library(gridExtra)
```

```{r}
current_file <- knitr::current_input()
wdir <- if (!is.null(current_file)) dirname(normalizePath(current_file)) else getwd()
ddir <- file.path(wdir, "data")


outdir <- file.path(wdir, "outputs/project2-public-data-viz")
dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
births.file <- file.path(ddir, "births_by_mothers_country_of_birth_2001_to_2024.csv")
geom.file <- file.path(ddir, "statistical-gis-boundaries-london.zip")
fb.file <- file.path(ddir, "fb-congress-data.csv")
election.file <- file.path(ddir, "2024_uk_election_results.csv")
const.file <- file.path(ddir, "mp_const.geojson")

stopifnot(file.exists(births.file))
stopifnot(file.exists(geom.file))
stopifnot(file.exists(fb.file))
stopifnot(file.exists(election.file))
stopifnot(file.exists(const.file))
```

2. London births data visualisation

2.1 Prepare London births and borough map data

```{r}
geom.dir <- file.path(ddir, "statistical-gis-boundaries-london")
if (!dir.exists(geom.dir)) {
  unzip(geom.file, exdir = ddir)
}

boro.file <- file.path(
  ddir,
  "statistical-gis-boundaries-london",
  "ESRI",
  "London_Borough_Excluding_MHW.shp"
)

stopifnot(file.exists(boro.file))

bf <- read_csv(births.file, show_col_types = FALSE)
gf <- read_sf(boro.file)

colnames(bf)
head(bf)
head(gf)
```

2.2 Total live births in London over time

```{r}
#计算每一年 London borough的total births
births_total_year <- bf |>
  filter(type == "London Borough") |>
  group_by(year) |>
  summarise(total_births_london = sum(total_births_all), .groups = "drop")

births_total_year

write_csv(births_total_year, file.path(outdir, "births_total_by_year.csv"))
```

```{r}
#live births柱状图
p_births_total <- births_total_year |>
  ggplot(aes(x = year, y = total_births_london)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title = "Total Live Births in London",
    subtitle = "London boroughs, 2001 to 2024",
    x = "Year",
    y = "Total live births"
  ) +
  theme_minimal()

p_births_total

ggsave(
  filename = file.path(outdir, "births_total_by_year.png"),
  plot = p_births_total,
  width = 9,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

2.3 UK-born and overseas-born mothers over time

```{r}
#按mother country of birth分组
births_mother_origin <- bf |>
  filter(type == "London Borough") |>
  group_by(year) |>
  summarise(
    UK_born_mothers = sum(total_births_uk_mothers),
    overseas_born_mothers = sum(total_births_overseas_mothers),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(UK_born_mothers, overseas_born_mothers),
    names_to = "mother_origin",
    values_to = "births"
  ) |>
  mutate(
    mother_origin = recode(
      mother_origin,
      "UK_born_mothers" = "UK-born mothers",
      "overseas_born_mothers" = "Overseas-born mothers"
    )
  )

births_mother_origin

write_csv(births_mother_origin, file.path(outdir, "births_by_mother_origin_year.csv"))
```

```{r}
# UK-born and overseas-born mothers 趋势图
p_births_origin_line <- births_mother_origin |>
  ggplot(aes(x = year, y = births, colour = mother_origin, group = mother_origin)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = label_comma()) +
  scale_x_continuous(breaks = seq(2001, 2024, 2)) +
  labs(
    title = "Live Births in London by Mother's Country of Birth",
    subtitle = "UK-born and overseas-born mothers, 2001 to 2024",
    x = "Year",
    y = "Live births",
    colour = "Mother's country of birth"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_births_origin_line

ggsave(
  filename = file.path(outdir, "births_mother_origin_line.png"),
  plot = p_births_origin_line,
  width = 9,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

```{r}
# UK-born and overseas-born mothers 堆叠柱状图
p_births_origin_bar <- births_mother_origin |>
  mutate(year = factor(year)) |>
  ggplot(aes(x = year, y = births, fill = mother_origin)) +
  geom_col() +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title = "Live Births in London by Mother's Country of Birth",
    subtitle = "Stacked bar chart, 2001 to 2024",
    x = "Year",
    y = "Live births",
    fill = "Mother's country of birth"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_births_origin_bar

ggsave(
  filename = file.path(outdir, "births_mother_origin_bar.png"),
  plot = p_births_origin_bar,
  width = 10,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

2.4 London borough maps

```{r}
#合并 borough geometry 和 births 
df_births_map <- gf |>
  inner_join(bf, by = c("GSS_CODE" = "gss_code")) |>
  mutate(prop_nonuk_mothers = total_births_overseas_mothers / total_births_all)

#对比
births_map_compare <- df_births_map |>
  filter(year %in% c(2001, 2024))
```

```{r}
# Total live births borough map
p_births_map_total <- births_map_compare |>
  ggplot() +
  geom_sf(aes(fill = total_births_all), color = "black", linewidth = 0.15) +
  scale_fill_gradient(
    "Total live births",
    low = "white",
    high = "magenta4",
    labels = label_comma()
  ) +
  labs(
    title = "Total Live Births in London",
    subtitle = "Borough-level comparison, 2001 and 2024"
  ) +
  facet_grid(cols = vars(year)) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    strip.text = element_text(size = 12)
  )

p_births_map_total

ggsave(
  filename = file.path(outdir, "births_total_map_2001_2024.png"),
  plot = p_births_map_total,
  width = 10,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

```{r}
# Proportion of births to overseas-born mothers borough map
p_births_map_nonuk <- births_map_compare |>
  ggplot() +
  geom_sf(aes(fill = prop_nonuk_mothers), color = "black", linewidth = 0.15) +
  scale_fill_gradient(
    "Proportion of\noverseas-born mothers",
    low = "white",
    high = "darkgreen",
    labels = label_percent()
  ) +
  labs(
    title = "Proportion of Births to Overseas-born Mothers",
    subtitle = "Borough-level comparison, 2001 and 2024"
  ) +
  facet_grid(cols = vars(year)) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    strip.text = element_text(size = 12)
  )

p_births_map_nonuk

ggsave(
  filename = file.path(outdir, "births_overseas_mothers_map_2001_2024.png"),
  plot = p_births_map_nonuk,
  width = 10,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

3. Facebook engagement visualisation

3.1 Prepare Facebook data

```{r}

fb <- read_csv(fb.file, show_col_types = FALSE)

#清理数据
fb_clean <- fb |>
  filter(!post_type %in% c("music", "note")) |>
  mutate(
    date = as.Date(date),
    gender = if_else(gender == "M", "Man", "Woman"),
    type = if_else(type == "rep", "House", "Senate")
  )

#基本检查
nrow(fb)
nrow(fb_clean)
head(fb_clean)

write_csv(fb_clean, file.path(outdir, "facebook_posts_cleaned.csv"))
```

3.2 Summary statistics for likes

```{r}
# likes summary statistics
fb_likes_summary <- fb_clean |>
  summarise(
    min_likes = min(likes_count, na.rm = TRUE),
    mean_likes = mean(likes_count, na.rm = TRUE),
    median_likes = median(likes_count, na.rm = TRUE),
    max_likes = max(likes_count, na.rm = TRUE),
    n_posts = n()
  )

fb_zero_likes <- fb_clean |>
  mutate(zero_likes = if_else(likes_count == 0, 1, 0)) |>
  summarise(zero_likes_percent = mean(zero_likes, na.rm = TRUE) * 100)

fb_likes_summary
fb_zero_likes

write_csv(fb_likes_summary, file.path(outdir, "facebook_likes_summary.csv"))
write_csv(fb_zero_likes, file.path(outdir, "facebook_zero_likes_percent.csv"))
```

3.3 Likes distribution

```{r}
p_fb_likes_hist <- fb_clean |>
  filter(likes_count > 0) |>
  ggplot(aes(x = likes_count)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  scale_x_log10("Number of likes", labels = comma) +
  scale_y_continuous("Number of posts", labels = comma) +
  labs(
    title = "Distribution of Facebook Likes",
    subtitle = "US Congress Facebook posts, log scale"
  ) +
  theme_minimal()

p_fb_likes_hist

ggsave(
  filename = file.path(outdir, "facebook_likes_histogram_log.png"),
  plot = p_fb_likes_hist,
  width = 8,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

```{r}
p_fb_likes_density_party <- fb_clean |>
  filter(likes_count > 0, party != "Independent") |>
  ggplot(aes(x = likes_count, colour = party)) +
  geom_density() +
  scale_x_log10("Number of likes", labels = comma) +
  labs(
    title = "Facebook Likes Distribution by Party",
    subtitle = "Density plot on log scale",
    colour = "Party"
  ) +
  theme_minimal()

p_fb_likes_density_party

ggsave(
  filename = file.path(outdir, "facebook_likes_density_by_party.png"),
  plot = p_fb_likes_density_party,
  width = 8,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

3.4 Engagement relationship: likes and comments

```{r}
# Scatter plot of likes and comments
p_fb_likes_comments <- fb_clean |>
  filter(likes_count > 0, comments_count > 0) |>
  ggplot(aes(x = likes_count, y = comments_count)) +
  geom_point(alpha = 0.25) +
  scale_x_log10("Likes count", labels = comma) +
  scale_y_log10("Comments count", labels = comma) +
  labs(
    title = "Relationship between Facebook Likes and Comments",
    subtitle = "Log scales used because engagement counts are skewed"
  ) +
  theme_minimal()

p_fb_likes_comments

ggsave(
  filename = file.path(outdir, "facebook_likes_comments_scatter.png"),
  plot = p_fb_likes_comments,
  width = 8,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

```{r}
# Likes and comments by post type
p_fb_likes_comments_type <- fb_clean |>
  filter(likes_count > 0, comments_count > 0) |>
  ggplot(aes(x = likes_count, y = comments_count)) +
  geom_point(alpha = 0.25) +
  scale_x_log10("Likes count", labels = comma) +
  scale_y_log10("Comments count", labels = comma) +
  facet_wrap(~post_type, nrow = 2) +
  labs(
    title = "Likes and Comments by Post Type",
    subtitle = "Facet plot by Facebook post type"
  ) +
  theme_minimal()

p_fb_likes_comments_type

ggsave(
  filename = file.path(outdir, "facebook_likes_comments_by_post_type.png"),
  plot = p_fb_likes_comments_type,
  width = 10,
  height = 6,
  dpi = 300,
  bg = "white"
)
```

3.5 Posts by party, gender, and post type

```{r}
# Monthly number of posts by party
fb_monthly_party <- fb_clean |>
  filter(party != "Independent") |>
  group_by(month = ceiling_date(date, "month"), party) |>
  summarise(posts = n(), .groups = "drop")

write_csv(fb_monthly_party, file.path(outdir, "facebook_monthly_posts_by_party.csv"))
```

```{r}
# Time series of monthly posts by party
p_fb_monthly_party <- fb_monthly_party |>
  ggplot(aes(x = month, y = posts, colour = party, group = party)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Monthly Facebook Posts by Party",
    x = "Month",
    y = "Number of posts",
    colour = "Party"
  ) +
  theme_minimal()

p_fb_monthly_party

ggsave(
  filename = file.path(outdir, "facebook_monthly_posts_by_party.png"),
  plot = p_fb_monthly_party,
  width = 8,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

```{r}
# Posts by post type and party
fb_post_type_party <- fb_clean |>
  filter(party != "Independent") |>
  group_by(post_type, party) |>
  summarise(posts = n(), .groups = "drop")

write_csv(fb_post_type_party, file.path(outdir, "facebook_posts_by_type_party.csv"))

p_fb_post_type_party <- fb_post_type_party |>
  ggplot(aes(x = party, y = post_type, fill = posts)) +
  geom_tile() +
  scale_fill_gradient("Number of posts", low = "white", high = "darkblue", labels = comma) +
  labs(
    title = "Facebook Post Types by Party",
    x = "Party",
    y = "Post type"
  ) +
  theme_minimal()

p_fb_post_type_party

ggsave(
  filename = file.path(outdir, "facebook_post_type_party_tile.png"),
  plot = p_fb_post_type_party,
  width = 7,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

```{r}
# Summary grid: post type, gender, average likes
p1 <- ggplot(fb_clean, aes(x = post_type)) +
  geom_bar() +
  scale_y_continuous("Number of posts", labels = comma) +
  scale_x_discrete("Type of post") +
  theme_minimal() +
  theme(axis.ticks.x = element_blank())

p2 <- ggplot(fb_clean, aes(x = gender)) +
  geom_bar() +
  scale_x_discrete("Gender of legislator") +
  scale_y_continuous("Number of posts", labels = comma) +
  theme_minimal()

p3 <- ggplot(fb_clean, aes(x = post_type, y = likes_count)) +
  stat_summary(fun = "mean", geom = "point", size = 4, shape = 15) +
  scale_y_continuous("Average likes count", labels = comma) +
  scale_x_discrete("Type of post") +
  theme_minimal()

p4 <- ggplot(fb_clean, aes(x = gender, y = likes_count)) +
  stat_summary(fun = "mean", geom = "point", size = 4, shape = 15) +
  scale_x_discrete("Gender of legislator") +
  scale_y_continuous("Average likes count", labels = comma) +
  theme_minimal()

# Save grid plot
png(file.path(outdir, "facebook_summary_grid.png"), width = 1800, height = 1400, res = 220)
grid.arrange(arrangeGrob(p1, p2, p3, p4, ncol = 2))
dev.off()
```

4. Optional extension: UK election results mapping

This section uses the 2024 UK election results and parliamentary constituency boundaries to create a constituency-level map.

```{r}
election_results <- read_csv(election.file, show_col_types = FALSE)
constituencies <- read_sf(const.file)

# 每个constituency票数最高的候选人或者政党就是 winner
election_winners <- election_results |>
  rename(
    ons_id = `ONS ID`,
    constituency_name = `Constituency name`,
    party_name = `Party name`,
    party_abbr = `Party abbreviation`,
    votes = Votes,
    share = Share
  ) |>
  group_by(ons_id, constituency_name) |>
  slice_max(order_by = votes, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(ons_id, constituency_name, party_name, party_abbr, votes, share)

# 统计各党seats
seat_summary <- election_winners |>
  count(party_abbr, party_name, name = "seats") |>
  arrange(desc(seats))

seat_summary

write_csv(election_winners, file.path(outdir, "uk_election_2024_winners.csv"))
write_csv(seat_summary, file.path(outdir, "uk_election_2024_seat_summary.csv"))
```

```{r}
#合并地图和 election winners
election_map <- constituencies |>
  inner_join(election_winners, by = c("PCON24CD" = "ons_id"))

#只突出主要政党，其他放进 Other
major_parties <- c("Lab", "Con", "LD", "SNP", "SF", "DUP", "RUK", "Green", "PC")

election_map <- election_map |>
  mutate(
    party_group = if_else(party_abbr %in% major_parties, party_abbr, "Other")
  )
```

```{r}
#Winning party map
p_election_winner_map <- election_map |>
  ggplot() +
  geom_sf(aes(fill = party_group), color = NA) +
  labs(
    title = "2024 UK General Election Results",
    subtitle = "Winning party by constituency",
    fill = "Winning party"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 11)
  )

p_election_winner_map

ggsave(
  filename = file.path(outdir, "uk_election_2024_winning_party_map.png"),
  plot = p_election_winner_map,
  width = 8,
  height = 10,
  dpi = 300,
  bg = "white"
)

```

```{r}
# Seat summary bar chart
p_election_seats <- seat_summary |>
  mutate(party_abbr = reorder(party_abbr, seats)) |>
  ggplot(aes(x = party_abbr, y = seats)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Seats Won by Party",
    subtitle = "2024 UK General Election",
    x = "Party",
    y = "Number of seats"
  ) +
  theme_minimal()

p_election_seats

ggsave(
  filename = file.path(outdir, "uk_election_2024_seats_by_party.png"),
  plot = p_election_seats,
  width = 8,
  height = 5,
  dpi = 300,
  bg = "white"
)
```

5. Output list

```{r}
list.files(outdir)
```
