---
title: "Project 3: Managing Social Media Data with Databases"
subtitle: "Using SQLite and SQL to Store, Query, and Join Congress Facebook Data"
author: "Zhao Zhanyu"
format:
  html:
    toc: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Project Overview

This project demonstrates how to use R, SQLite, and SQL to store, manage, query, and join a large social media dataset. The dataset contains Facebook posts from members of the U.S. Congress in 2017. The workflow stores account-level metadata and post-level Facebook data in a local SQLite database, then uses SQL queries and joins to prepare the data for communication analysis.

0. Setup

```{r}
required_packages <- c("tidyverse", "DBI", "RSQLite", "janitor", "scales")

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
}

library(tidyverse)
library(DBI)
library(RSQLite)
library(janitor)
library(scales)
```

```{r}
current_file <- knitr::current_input()
wdir <- if (!is.null(current_file)) dirname(normalizePath(current_file)) else getwd()

ddir <- file.path(wdir, "data")
outdir <- file.path(wdir, "outputs/project3-social-media-database")

dir.create(ddir, recursive = TRUE, showWarnings = FALSE)
dir.create(outdir, recursive = TRUE, showWarnings = FALSE)

zip_candidates <- c(
  file.path(ddir, "facebook-data.zip"),
  file.path(wdir, "facebook-data.zip")
)

fb_zip <- zip_candidates[file.exists(zip_candidates)][1]

fb_dir <- file.path(ddir, "facebook-data")

# The zip file may extract files directly into fb_dir, or into a nested facebook-data folder.
if (!file.exists(file.path(fb_dir, "congress-facebook-2017.csv")) &&
    file.exists(file.path(fb_dir, "facebook-data", "congress-facebook-2017.csv"))) {
  fb_dir <- file.path(fb_dir, "facebook-data")
}

congress_file <- file.path(fb_dir, "congress-facebook-2017.csv")
posts_dir <- file.path(fb_dir, "posts")
```

1. Create a SQLite Database

```{r}
db_path <- file.path(outdir, "facebook-db.sqlite")

if (file.exists(db_path)) {
  file.remove(db_path)
}

#Create and connect to the SQLite database.
db <- dbConnect(RSQLite::SQLite(), db_path)
```

2. Add Congress Metadata Table

```{r}
congress <- read_csv(congress_file, show_col_types = FALSE) |>
  clean_names()

dbWriteTable(db, "congress", congress, overwrite = TRUE)

dbListFields(db, "congress")

tibble(field = dbListFields(db, "congress")) |>
  write_csv(file.path(outdir, "congress_table_fields.csv"))
```

3. Add Facebook Post Data Table

```{r}
post_files <- list.files(posts_dir, pattern = "\\.csv$", full.names = TRUE)

length(post_files)

# Read each post file and append it to the posts table.
# All columns are read as character first, which is safer for large social media files.
for (f in post_files) {
  post_data <- read_csv(
    f,
    show_col_types = FALSE,
    locale = locale(encoding = "latin1"),
    col_types = cols(.default = col_character())
  ) |>
    clean_names()
  
  dbWriteTable(db, "posts", post_data, append = TRUE)
  
  rm(post_data)
}

# Check table fields.
dbListFields(db, "posts")

tibble(field = dbListFields(db, "posts")) |>
  write_csv(file.path(outdir, "posts_table_fields.csv"))
```

4. Check Database Tables

```{r}
dbListTables(db)

# Count rows in each table.
table_counts <- tibble(
  table = c("congress", "posts"),
  rows = c(
    dbGetQuery(db, "SELECT COUNT(*) AS n FROM congress")$n,
    dbGetQuery(db, "SELECT COUNT(*) AS n FROM posts")$n
  )
)

table_counts

write_csv(table_counts, file.path(outdir, "database_table_counts.csv"))
```

5. Basic SQL Queries

```{r}
congress_head <- dbGetQuery(db, "
  SELECT *
  FROM congress
  LIMIT 10
")

congress_head

write_csv(congress_head, file.path(outdir, "congress_head.csv"))
```

```{r}
# Select only selected columns.
congress_name_party <- dbGetQuery(db, "
  SELECT name, party, type, screen_name
  FROM congress
  LIMIT 20
")

congress_name_party

write_csv(congress_name_party, file.path(outdir, "congress_name_party_sample.csv"))
```

```{r}
republican_members <- dbGetQuery(db, "
  SELECT name, party, type, screen_name
  FROM congress
  WHERE party = 'Republican'
  LIMIT 20
")

republican_members

write_csv(republican_members, file.path(outdir, "republican_members_sample.csv"))
```

6. Join Congress Metadata with Facebook Posts

```{r}
# Join the congress metadata table with the posts table using screen_name.
# CAST() is used because post metrics were first read as character columns.
joined_sample <- dbGetQuery(db, "
  SELECT
    c.name,
    c.party,
    c.type AS member_type,
    p.screen_name,
    p.date,
    p.type AS post_type,
    CAST(p.likes_count AS INTEGER) AS likes_count,
    CAST(p.comments_count AS INTEGER) AS comments_count,
    CAST(p.shares_count AS INTEGER) AS shares_count,
    p.message
  FROM posts AS p
  INNER JOIN congress AS c
    ON p.screen_name = c.screen_name
  LIMIT 50
")

joined_sample

write_csv(joined_sample, file.path(outdir, "joined_posts_congress_sample.csv"))
```

7. SQL Summary: Engagement by Party

```{r}
party_engagement <- dbGetQuery(db, "
  SELECT
    c.party,
    COUNT(*) AS n_posts,
    SUM(CAST(p.likes_count AS INTEGER)) AS total_likes,
    ROUND(AVG(CAST(p.likes_count AS INTEGER)), 2) AS avg_likes,
    SUM(CAST(p.comments_count AS INTEGER)) AS total_comments,
    ROUND(AVG(CAST(p.comments_count AS INTEGER)), 2) AS avg_comments,
    SUM(CAST(p.shares_count AS INTEGER)) AS total_shares,
    ROUND(AVG(CAST(p.shares_count AS INTEGER)), 2) AS avg_shares
  FROM posts AS p
  INNER JOIN congress AS c
    ON p.screen_name = c.screen_name
  GROUP BY c.party
  ORDER BY n_posts DESC
")

party_engagement

write_csv(party_engagement, file.path(outdir, "party_engagement_summary.csv"))
```

8. SQL Summary: Top Accounts by Total Likes

```{r}
top_accounts <- dbGetQuery(db, "
  SELECT
    c.name,
    c.party,
    c.type AS member_type,
    p.screen_name,
    COUNT(*) AS n_posts,
    SUM(CAST(p.likes_count AS INTEGER)) AS total_likes,
    ROUND(AVG(CAST(p.likes_count AS INTEGER)), 2) AS avg_likes,
    SUM(CAST(p.comments_count AS INTEGER)) AS total_comments,
    SUM(CAST(p.shares_count AS INTEGER)) AS total_shares
  FROM posts AS p
  INNER JOIN congress AS c
    ON p.screen_name = c.screen_name
  GROUP BY c.name, c.party, c.type, p.screen_name
  ORDER BY total_likes DESC
  LIMIT 20
")

top_accounts

write_csv(top_accounts, file.path(outdir, "top_accounts_by_total_likes.csv"))
```

9. Check Unmatched Congress Accounts

```{r}
unmatched_congress <- dbGetQuery(db, "
  SELECT DISTINCT
    c.screen_name,
    c.name,
    c.party,
    c.type
  FROM congress AS c
  LEFT JOIN posts AS p
    ON c.screen_name = p.screen_name
  WHERE p.screen_name IS NULL
  ORDER BY c.party, c.name
")

unmatched_congress

write_csv(unmatched_congress, file.path(outdir, "unmatched_congress_accounts.csv"))
```

10. Visualise SQL Outputs

```{r}
party_plot <- party_engagement |>
  ggplot(aes(x = reorder(party, n_posts), y = n_posts)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of Facebook Posts by Party",
    x = "Party",
    y = "Number of posts"
  ) +
  theme_minimal()

party_plot

ggsave(
  filename = file.path(outdir, "posts_by_party.png"),
  plot = party_plot,
  width = 8,
  height = 5,
  dpi = 300
)
```

```{r}
top_accounts_plot <- top_accounts |>
  mutate(name = fct_reorder(name, total_likes)) |>
  ggplot(aes(x = name, y = total_likes)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 20 Congress Facebook Accounts by Total Likes",
    x = "Account",
    y = "Total likes"
  ) +
  theme_minimal()

top_accounts_plot

ggsave(
  filename = file.path(outdir, "top_accounts_by_total_likes.png"),
  plot = top_accounts_plot,
  width = 9,
  height = 6,
  dpi = 300
)
```

